Supporting information for ``How environmental drivers of synchrony interact''

### 2022-01-16 Paper/SupMat.Rnw
continuting to write and edit

Now, having specified the noise process $(\epsilon^{(1)},\epsilon^{(2)})$, we derive the quantities 
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$.
For $a$ and $b$ equal to $1$ or $2$, the formula 
$S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = \sum_{u=-\infty}^\infty c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)
\exp(-2 \pi \iota f u)$, where $c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$,
follows from definitions of the spectrum and cross spectrum, given, for instance, in 2.5.2 on p. 23 of
\cite{Brillinger2001}. The quantity 
$c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$ is independent of 
$t$ for any stationary process $(\epsilon^{(1)},\epsilon^{(2)})$, and in particular for the temporally iid process that 
we assumed to operate here. Furthermore, because 
$\epsilon_i^{(a)}(t+u)$ and $\epsilon_j^{(b)}(t)$ were assumed independent for $u \neq 0$, we have 
\begin{equation}
S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(0) =
\begin{cases}
1 & \text{if}\ a=b\ \text{and}\ i=j \\
v_{11} & \text{if}\ a=b=1\ \text{and}\ i \neq j \\
v_{22} & \text{if}\ a=b=2\ \text{and}\ i \neq j \\
v_{12} & \text{if}\ a \neq b.
\end{cases}
\end{equation}
Thus the spectra and cross spectra in this example are constant. It
is then easy to see that $\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$, based on the definitions of these 
quantities given previously. For theoretical case studies A and B in the main text, we used
$N=\Sexpr{N_A}$, $v_{11}=v{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.

@@

@@

Floating surface canopy of kelp was measured by remote sensing, with details and
validation reported by \cite{Cavanaugh2011} and \cite{Bell2020, Bell2021}. Multispectral Landsat
imagery of $30$m resolution was used. Quarterly mean biomass values and spatial extent of canopy  
within $500$m coastline segments (henceforth called ``loactions'')
were computed by averaging. Quarters were: Q1, January-March; Q2, April-June; 
Q3, July-September; Q4, October-December. 
Data were thusly available from $\Sexpr{dim(kelp_SB)[1]}$ locations spanning 
approximately from Monterey Bay to San Diego. 
We then chose locations for analysis for which kelp was ``persistent'', in this 
context meaning absent or missing data for no more than three years; this was to focus on 
population dynamics of kelp rather than extinction-recolonization dynamics, which occur in some 
locations and which can relate to sediment dynamics on the ocean floor and other factors
not currently of primary interest. Of the original $\Sexpr{dim(kelp_SB)[1]}$ locations,
$\Sexpr{dim(kelp_MI)[1]}$ ($\Sexpr{100*round(dim(kelp_MI)[1]/dim(kelp_SB)[1],2)}$\%) 
were deemed persistent according to this criterion. Missing values in the remaining 
time series were then filled with seasonal medians. The biomass and spatial extent time series
for each location were then used together to produce one time series estimating biomass
density per unit useable habitat at the site, as follows: every value of 
the biomass time series was divided by the $90^{\text{th}}$ quantile of the distribution
of the spatial extent time series at the site. The $90^{\text{th}}$ quantile was used instead
of the maximum value of the spatial extent time series because maximum statistics have high variance. 
This procedure served to normalize total biomass by an estimate
of total useable habitat at a site. Kelp establish on rocky substrate, and areas with sandy or
sedimented substrate are not useable kelp habitat. All time series were then linearly detrended,
and only those time series were utilized which were located within the three regions pictured
in Fig. \ref{MT-fig:map} in the main text. Regions were selected to represent distinct dynamical
patterns known to occur in central and southern CA, and also to have another the same number
of time series per selected region.

>>>>

Now, having specified the noise process $(\epsilon^{(1)},\epsilon^{(2)})$, we derive the quantities 
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$.
For $a$ and $b$ equal to $1$ or $2$, the formula 
$S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = \sum_{u=-\infty}^\infty c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)
\exp(-2 \pi \iota f u)$, where $c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$,
follows from definitions of the spectrum and cross spectrum, given, for instance, in 2.5.2 on p. 23 of
\cite{Brillinger2001}. The quantity 
$c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$ is independent of 
$t$ for any stationary process $(\epsilon^{(1)},\epsilon^{(2)})$, and in particular for the temporally iid process that 
we assumed to operate here. Furthermore, because 
$\epsilon_i^{(a)}(t+u)$ and $\epsilon_j^{(b)}(t)$ were assumed independent for $u \neq 0$, we have 
\begin{equation}
S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(0) =
\begin{cases}
1 & \text{if}\ a=b\ \text{and}\ i=j \\
v_{11} & \text{if}\ a=b=1\ \text{and}\ i \neq j \\
v_{22} & \text{if}\ a=b=2\ \text{and}\ i \neq j \\
v_{12} & \text{if}\ a \neq b.
\end{cases}
\end{equation}
Thus the spectra and cross spectra in this example are constant. It
is then easy to see that $\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$, based on the definitions of these 
quantities given previously. For theoretical case studies A and B in the main text, we used
$N=\Sexpr{N_A}$, $v_{11}=v_{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.

@@

Floating surface canopy of kelp was measured by remote sensing, with details and
validation reported by \cite{Cavanaugh2011} and \cite{Bell2020, Bell2021}. Multispectral Landsat
imagery of $30$m resolution was used. Quarterly mean biomass values and spatial extent of canopy  
within $500$m coastline segments (henceforth called ``locations'')
were computed by averaging. Quarters were: Q1, January-March; Q2, April-June; 
Q3, July-September; Q4, October-December. 
Data were thusly available from $\Sexpr{dim(kelp_SB)[1]}$ locations spanning 
approximately from Monterey Bay to San Diego. 
We then chose locations for analysis for which kelp was ``persistent'', in this 
context meaning absent or missing data for no more than three years; this was to focus on 
population dynamics of kelp rather than extinction-recolonization dynamics, which occur in some 
locations and which can relate to sediment dynamics on the ocean floor and other factors
not currently of primary interest. Of the original $\Sexpr{dim(kelp_SB)[1]}$ locations,
$\Sexpr{dim(kelp_MI)[1]}$ ($\Sexpr{100*round(dim(kelp_MI)[1]/dim(kelp_SB)[1],2)}$\%) 
were deemed persistent according to this criterion. Missing values in the remaining 
time series were then filled with seasonal medians. The biomass and spatial extent time series
for each location were then used together to produce one time series estimating biomass
density per unit useable habitat at the site, as follows: every value of 
the biomass time series was divided by the $90^{\text{th}}$ quantile of the distribution
of the spatial extent time series at the site. The $90^{\text{th}}$ quantile was used instead
of the maximum value of the spatial extent time series because maximum statistics have high variance. 
This procedure served to normalize total biomass by an estimate
of total useable habitat at a site. Kelp establish on rocky substrate, and areas with sandy or
sedimented substrate are not useable kelp habitat. All time series were then linearly detrended,
and only those time series were utilized which were located within the three regions pictured
in Fig. \ref{MT-fig:map} in the main text. Regions were selected to represent distinct dynamical
patterns known to occur in central and southern CA, and also to have another the same number
of time series per selected region.

@@
